---
title: "spatial examples"
author: "Alex Bushby, Aleks Jovic, Ben Bolker"
---

```{r pkgs,message=FALSE}
library(ggmap)
library(tidyverse)
library(viridis)
library(sf)
load("data/googlemaps.rda")
```

# crime

The `crime` dataset contains crime reports for Houston from January-August 2010, geocoded with Google Maps.

```{r showcrime}
tibble(crime)
```
Lots of useful info: dates, types of crimes, locations by type of place, locations by street, locations by longitude/latitude ...

First, let's get an overview of the crimes on the map. The `qmplot` is often recommended as a quick way to do mapping (but we will switch to another approach shortly). We put in longitude (`lon`) and latitude (`lat`) for the `x` and `y` parameters and specify `crime` as the data set. This plots all of the crimes in the database.

(Example adapted from [here](https://github.com/tidyverse/ggplot2/wiki/crime-in-downtown-houston,-texas-:-combining-ggplot2-and-google-maps).)

```{r qmplot,message=FALSE}
q1 <- qmplot(lon, lat, data = crime,
             maptype = "toner-lite",
             ## for q* plots need to use I() for direct specifications
             colour = I("red"),
             size = I(0.9),
             alpha=I(.3))
```

A slightly slower but safer method is to get the map first, then combine it with the data. This way we can retrieve the map and store it; this is both more efficient if we're going to make a bunch of plots with the same map (almost inevitable if we're polishing a data visualization), and safer (in case the server goes down/network connection is lost/etc.).

```{r houston1}
## utility function: extract appropriate components for retrieving a Stamen/OSM map
get_mapzone <- function(data, latcol="lat", loncol="lon") {
  lon <- na.omit(data[[loncol]])
  lat <- na.omit(data[[latcol]])
  return(c(left=min(lon),right=max(lon),bottom=min(lat),top=max(lat)))
}
houston_mz <- get_mapzone(crime) ## find boundaries
## retrieve map
houston_map1 <- get_stamenmap(houston_mz,zoom=7, maptype="toner-lite")
```

Now we can combine the map with the data set:

```{r houston2}
(ggmap(houston_map1)
  + geom_point(data=crime,colour="red",size=0.9,alpha=0.3)
)
```

This graph is not very good, in particular because of the 


Reduce crime to violent crimes in downtown Houston:

```{r violent_crime}
violent_crime <-
  (crime
    %>% filter(
          !offense %in% c("auto theft", "theft", "burglary"),
          -95.39681 <= lon & lon <= -95.34188,
          29.73631 <= lat & lat <=  29.78400
        )
    %>% mutate(
          ## drops unused levels, mitigates downstream errors
          offense = fct_drop(offense), 
          offense = fct_relevel(offense,
               c("robbery", "aggravated assault", "rape", "murder")
               )
        )
  )
```


```{r vc1}
houston_map2 <- get_stamenmap(get_mapzone(violent_crime), maptype="toner-lite",
                              zoom=14)
(ggmap(houston_map2)
  + geom_point(data = violent_crime, colour = "red",size = 0.9, alpha=.3)
)
```

Re-use the map, plotting as density contours instead (i.e., transform the points to a density field, then summarize the density field as a set of contours)

```{r vc2}
(ggmap(houston_map2)
  + geom_density2d(data = violent_crime, aes(x=lon,y=lat), col="red")
)
```

Or do 2-D (square) binning, with a custom gradient (and log-scaled breaks):

```{r vc3}
(ggmap(houston_map2)
  + geom_bin2d(data = violent_crime, alpha=0.5)
  + scale_fill_gradient(low="#F0F0FF", high="#131393",
                        trans=scales::log10_trans())
)
```

The high-density square on Smith Street is messing up our ability to see much else.

## Heatmap

To make the contour map more useful, we can assign a gradient (using a "polygon" rather than a "density_2d" geom with `stat_density_2d`) . Let’s look at the robberies:

```{r vc4}
## "background" rather than "lite" map
houston_map3 <- get_stamenmap(get_mapzone(violent_crime),
                              maptype="toner-background",
                              zoom=14)
```

```{r vc5}
robbery <- violent_crime %>% filter(offense=='robbery')
(ggmap(houston_map3)
  + stat_density_2d(data=robbery,
                    aes(fill = ..level..), geom = "polygon",
                    alpha = .35, colour = NA)
  + scale_fill_gradient2("Robbery\nheatmap",
                         low = "white", mid = "yellow",
                         high = "red", midpoint = 650)
)
```

We can do all the other `ggplot` stuff, like faceting:

```{r facet1}
(ggmap(houston_map3,darken=c(0.9,"white")) ## fade map layer
  + geom_point(data = violent_crime, aes(colour = offense, size=offense))
  + facet_wrap(~ offense)
  + scale_colour_brewer(palette="Dark2")
  + scale_size_manual(values=c(1,1,2,3)) ## adjust point sizes for visibility
  + theme(legend.position="none")
)
```

```{r weekday_facets}
hm <- ggmap(houston_map3, base_layer = ggplot(aes(x = lon, y = lat),
                                              data = violent_crime),
            darken=c(0.9,"white"))
(hm
  + stat_density2d(aes(fill = ..level.., alpha = ..level..),
                bins = 5, geom = "polygon")
  + scale_fill_gradient(low = "black", high = "red")
  + facet_wrap(~ day)
  + theme(legend.position="none")
)
```

(there are some contouring artifacts here I don't understand ...)

This is pretty but maybe not useful ...

```{r crime_density}
(hm
  + stat_density2d(aes(x = lon, y = lat, fill = offense, alpha = ..level..),
                   bins = 5, geom = "polygon")
  + scale_fill_brewer(palette="Dark2")
  
)
```

There is not a super-easy way to code scales (we'd like 29.75 to be $29^\circ 45' N$, e.g. see [here](https://stackoverflow.com/questions/33302424/format-latitude-and-longitude-axis-labels-in-ggplot) for a partial solution), but maybe we don't even need them on a map?

# Canadian electoral ridins

From [this github site](https://github.com/paleolimbot/blogdown-site/tree/master/content/post/2019-10-21-ridings):

The riding boundaries come in shapefile format, which is handled by a one-liner to `read_sf()`. I'm going to simplify the boundaries a bit to speed up the plotting time. The `dTolderance` argument is in map units, and it took some experimenting to settle on the number 100.

```{r}
ridings <- (read_sf("boundaries_2015_shp_en/FED_CA_2_2_ENG.shp")
  %>% st_simplify(dTolerance = 100)
)
ridings

